Spatial Reasoning and Statistical Graphics

### 2014-11-10 Paper/Draft1/TestingVisualAptitude.Rnw
Reorder boxplot levels so that "boxplot" is located between the two relevant experiments.

<<lineup-types,echo=F, fig.cap="Plot of scaled scores for different types of lineups\\label{fig:plottypesdensity}", fig.width=3, fig.height=6, out.width=".9\\linewidth", warning=F, message=F>>=
lineup.type.sum2$plot.type <- str_replace(lineup.type.sum2$plot.type, "rotated", "rot")
lineup.type.sum2$plot.type <- str_replace(as.character(lineup.type.sum2$plot.type), "error", "err")
lineup.type.sum2$plot.type <- factor(lineup.type.sum2$plot.type, 
                                     levels=c("boxplot", "jitter.boxplot", "violinplot", "bee.swarm.boxplot", 
                                              "scatterplot", "density", "histogram", 
                                              "qq.line", "qq.line.err", "qq.line.err.rot"), 
                                     labels=c("Boxplot", "Boxplot + \nOverlaid Data", "Violin Plot", "Bee Swarm Boxplot",
                                              "Scatterplot", "Density Plot", "Overlaid Histogram", 
                                              "QQ plot + Line", "QQ plot +\nAcceptance Band", 
                                              "Rotated QQ plot +\nAcceptance Band")
                                     )
qplot(data=subset(lineup.type.sum2, !is.na(plot.type)& testtype=="lineup"), 
      x=score, y=..scaled.., colour=factor(testnum), geom="density", cut=4) + 
  facet_wrap(~plot.type, ncol=2) + 
  xlab("Proportion of Lineups Correct") + 
  ylab("Density") + 
  geom_vline(aes(xintercept=0), linetype=3) + 
  scale_colour_discrete("Lineup Test Number") + 
  scale_x_continuous(breaks=c(-.33, 0, .33, .67, 1)) +
  scale_y_continuous(breaks=c(0, .5, 1)) +
  theme_bw() + 
  theme(legend.position="bottom", legend.direction="horizontal", 
        axis.text.x=element_text(angle=30, hjust=1),
        plot.margin=unit(c(0,0,0,0),"mm"), 
        axis.ticks.y=element_blank(), axis.text.y=element_blank())
@
\newtext{
Figure \ref{fig:plottypesdensity} compares performance on each different type of plot. As two different lineup tasks utilized boxplots to test different qualities of the distribution of data (outliers vs. difference in medians), different tasks are shown as different colors, so that accuracy on tasks which are shown in blue can be compared to other blue density curves. 
}

\newtext{
Figure \ref{fig:scatterplottypes} shows the association between scaled score on each type of lineup and score on the visual reasoning tests. Sample size for each plot type is fairly small - between 5 and 10 plots per individual, so there is low power for systematic inference, but we can establish that the card rotation task is much more significantly associated with the QQ plots tasks compared to the other tasks. In addition, rotated qq plots seem to be much more asssociated with the paper folding task scores than other qq plot tasks; this may be because they require more visual manipulation than other qq plots?
}
<<lineup-types-scores,echo=F, fig.cap="Plot of scaled lineup scores by aptitude test scores.\\label{fig:scatterplottypes}", fig.width=6, fig.height=12, out.width="\\linewidth", fig.env="figure", warning=F, message=F>>=
lineup.type.sum3 <- dcast(lineup.type.sum2, id + testtype ~ plot.type, value.var="score", fun.aggregate = mean)
other.scores <- subset(lineup.type.sum3, testtype!="lineup")[,c(1,2,13)]
names(other.scores)[3] <- "testscore"
lineup.scores <- melt(subset(lineup.type.sum3, testtype=="lineup")[,c(1,3:12)], id.vars = 1)
names(lineup.scores)[2:3] <- c("lineup.type", "lineup.pct")
lineup.type.sum3 <- merge(other.scores, lineup.scores, all.x=T, all.y=T)
lineup.type.sum3$lineup.type <- factor(lineup.type.sum3$lineup.type, 
                                     levels=c("Boxplot", "Boxplot + \nOverlaid Data", "Violin Plot", "Bee Swarm Boxplot",
                                              "Scatterplot", "Density Plot", "Overlaid Histogram", 
                                              "QQ plot + Line", "QQ plot +\nAcceptance Band", 
                                              "Rotated QQ plot +\nAcceptance Band"), 
                                     labels=c("Boxplot", "Boxplot\n + Points", "Violin Plot", "Bee Swarm \nBoxplot",
                                              "Scatter\nPlot", "Density\nPlot", "Overlaid\nHistogram", 
                                              "QQ plot\n+ Line", "QQ plot\n+ Bands", 
                                              "Rotated QQ\nplot + Bands")
                                     )

>>>>

<<lineup-types,echo=F, fig.cap="Plot of scaled scores for different types of lineups\\label{fig:plottypesdensity}", fig.width=3, fig.height=6, out.width=".9\\linewidth", warning=F, message=F>>=
lineup.type.sum2$plot.type <- str_replace(lineup.type.sum2$plot.type, "rotated", "rot")
lineup.type.sum2$plot.type <- str_replace(as.character(lineup.type.sum2$plot.type), "error", "err")
lineup.type.sum2$plot.type <- factor(lineup.type.sum2$plot.type, 
                                     levels=c("violinplot", "jitter.boxplot", "bee.swarm.boxplot", "boxplot", 
                                              "scatterplot", "density", "histogram", 
                                              "qq.line", "qq.line.err", "qq.line.err.rot"), 
                                     labels=c("Violin Plot", "Boxplot + \nOverlaid Data", "Bee Swarm Boxplot", "Boxplot", 
                                              "Scatterplot", "Density Plot", "Overlaid Histogram", 
                                              "QQ plot + Line", "QQ plot +\nAcceptance Band", 
                                              "Rotated QQ plot +\nAcceptance Band")
                                     )
qplot(data=subset(lineup.type.sum2, !is.na(plot.type)& testtype=="lineup"), 
      x=score, y=..scaled.., colour=factor(testnum), geom="density", cut=4) + 
  facet_wrap(~plot.type, ncol=2) + 
  xlab("Proportion of Lineups Correct") + 
  ylab("Density") + 
  geom_vline(aes(xintercept=0), linetype=3) + 
  scale_colour_discrete("Lineup Test Number") + 
  scale_x_continuous(breaks=c(-.33, 0, .33, .67, 1)) +
  scale_y_continuous(breaks=c(0, .5, 1)) +
  theme_bw() + 
  theme(legend.position="bottom", legend.direction="horizontal", 
        axis.text.x=element_text(angle=30, hjust=1),
        plot.margin=unit(c(0,0,0,0),"mm"), 
        axis.ticks.y=element_blank(), axis.text.y=element_blank())
@
\newtext{
Figure \ref{fig:plottypesdensity} compares performance on each different type of plot. As two different lineup tasks utilized boxplots to test different qualities of the distribution of data (outliers vs. difference in medians), different tasks are shown as different colors, so that accuracy on tasks which are shown in blue can be compared to other blue density curves. 
}

\newtext{
Figure \ref{fig:scatterplottypes} shows the association between scaled score on each type of lineup and score on the visual reasoning tests. Sample size for each plot type is fairly small - between 5 and 10 plots per individual, so there is low power for systematic inference, but we can establish that the card rotation task is much more significantly associated with the QQ plots tasks compared to the other tasks. In addition, rotated qq plots seem to be much more asssociated with the paper folding task scores than other qq plot tasks; this may be because they require more visual manipulation than other qq plots?
}
<<lineup-types-scores,echo=F, fig.cap="Plot of scaled lineup scores by aptitude test scores.\\label{fig:scatterplottypes}", fig.width=6, fig.height=12, out.width="\\linewidth", fig.env="figure", warning=F, message=F>>=
lineup.type.sum3 <- dcast(lineup.type.sum2, id + testtype ~ plot.type, value.var="score", fun.aggregate = mean)
other.scores <- subset(lineup.type.sum3, testtype!="lineup")[,c(1,2,13)]
names(other.scores)[3] <- "testscore"
lineup.scores <- melt(subset(lineup.type.sum3, testtype=="lineup")[,c(1,3:12)], id.vars = 1)
names(lineup.scores)[2:3] <- c("lineup.type", "lineup.pct")
lineup.type.sum3 <- merge(other.scores, lineup.scores, all.x=T, all.y=T)
lineup.type.sum3$lineup.type <- factor(lineup.type.sum3$lineup.type, 
                                     levels=c("Violin Plot", "Boxplot + \nOverlaid Data", "Bee Swarm Boxplot", "Boxplot", 
                                              "Scatterplot", "Density Plot", "Overlaid Histogram", 
                                              "QQ plot + Line", "QQ plot +\nAcceptance Band", 
                                              "Rotated QQ plot +\nAcceptance Band"), 
                                     labels=c("Violin Plot", "Boxplot\n + Points", "Bee Swarm \nBoxplot", "Boxplot", 
                                              "Scatter\nPlot", "Density\nPlot", "Overlaid\nHistogram", 
                                              "QQ plot\n+ Line", "QQ plot\n+ Bands", 
                                              "Rotated QQ\nplot + Bands")
                                     )
